Figures to insert in text

### 2018-03-06 Chapters/Chapter02.tex
sdm

The objective here is twofold: (i) a single command will install the framework, and (ii) another single command will run (and display) the desired figures obtained the simulation. This will allow potential users to become familiarized with the system and its underlying code with barely any learning curve beyond scientific python and jupyter notebooks.

@@

\subsection{Some initial anomalous results}

@@

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
\caption{Results generated by the framework similar from Kanerva's original Table 7.2. It was a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
\label{sdm-100w-table-7-2}}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_01-05.png}}
  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_15_3_45_6.png}}

  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=2.9in]{./images02/new-images/z_0_1.png}}
  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=2.9in]{./images02/new-images/iter_z_all_2.png}}

  \caption{(a) and (b) show the behavior of a single read; (c) and (d) present the effects of 6 iterative reads. As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.  Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.  It seems that Kanerva may not have considered that a single read would only `clean' a small number of dimensions \emph{after the critical distance}. What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
  \label{fig:murillo-generalization-experiments}
\end{figure}

Physicist Paulo Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values. The results, however, have not yielded performance improvements.  Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads. This is shown in Figure \ref{fig:murillo-generalization-experiments}.

@@

\begin{figure}[h!]
  \centering
  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}

  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}

  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}

  \caption{Computing the amount of information of a signal to each hard location in its access radius. (a) entirety of the space; (b) region of interest; (c) Fast computation is possible through a stepwise function. }
  \label{fig:info-theory-hypothesis}
\end{figure}

\begin{figure}[h!]
  \centering
  \subfloat[$w(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}

  \subfloat[$w(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}

  \subfloat[stepwise $\left \lfloor{w(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}

  \caption{Computing the sum of low-likelihood signals. (a) entirety of the space; (b) region of interest; (c) Fast computation through a stepwise function. }
  \label{fig:info-theory-sum-hypothesis}
\end{figure}

@@

My advisor, Alexandre Linhares, has proposed another read operation: an information-theoretical weighted reading. In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address. The logic behind it is to vary the importance of each hard-location inside the circle.  It is only natural that one encodes an item in near hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.  Closer hard locations have lower probabilies and therefore should encode more information.

@@

The results can be seen in Figure \ref{fig:info-theory-experiments} and seem promising. It seems that the critical distance increases by a number of bits.  Note that 10 additional bits imply an attractor $2^{10}$ of the size of the original. Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.

@@

Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.  More recently, he suggested the use of 10,000 dimensions.

@@

$\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.

\subsection{Docker image and jupyter notebooks}

We have generated a Docker image, which makes it even easier to explore the framework. After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed. The simulations run in this thesis are promptly available to be re-executed and explored. We invite readers to take a look and explore a little bit.

>>>>

\section{Distance between random bitstrings}

As showed by \citet{Kanerva1988}, the distance between two bitstrings follows a binomial distribution with mean $\mu = n/2$ and standard deviation $\sigma = \sqrt{n}/2$. For large values of $n$, it may be approximated by a normal distribution with the same mean and standard deviation.

In order to validate our random bitstring generation algorithm, we have calculated 10,000 distances between two random bitstrings with $n=1,000$ bits. In total, 20,000 random bitstrings have been generated during the simulation. The code is available in the ``Distance between bitstrings'' notebook \citep{sdmframework}.

In figure \ref{fig:validation-distance}, we can notice that the theoretical model and the simulation matches. Hence, it seems the random bitstring generation algorithm works properly.

This also validates the algorithm used to calculate the distance between two bitstrings. In this simulation, we have used the built-in \_\_popcnt() function.

\begin{figure}[!htb]
  \centering
  \subfloat[Full histogram ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-full.png}}
  \subfloat[{Zoom in the interval $[400, 600]$} ]{\includegraphics[width=0.5\textwidth]{./images02/new-images/bs-hist-zoom.png}}

  \caption{Histogram of 10,000 distances between two random bitstrings with 1,000 bits. The curve in red is the theoretical normal distribution with $\mu = 500$ and $\sigma = \sqrt{500}/2$.}
  \label{fig:validation-distance}
\end{figure}

@@

\section{Some initial anomalous results}

@@

\begin{figure}[h]
\centering\includegraphics[width=0.8\textwidth]{images02/sdm-100w-table-7-2.png}
\caption{Results generated by the framework similar from Kanerva's original Table 7.2. Here we have a 1,000 bit, 1,000,000 hard-location SDM with exactly 100 random bitstrings written into it.
\label{sdm-100w-table-7-2}}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \subfloat[$z \in \{.1, .2, .3, .4, .5, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_01-05.png}}

  \subfloat[$z \in \{1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_15_3_45_6.png}}

  \caption{(a) and (b) show the behavior of a single read; As stated previously, we can see a deterioration of convergence, with lower critical distance as $z>1$.  Another observation can be made here, concerning the discrepancy of Kanerva's Fig 7.3 and our data.  It seems that Kanerva may not have considered that a single read would only `clean' a small number of dimensions \emph{after the critical distance}. What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
  \label{fig:murillo-generalization-experiments}
\end{figure}

\begin{figure}[h!]
  \centering
  \subfloat[$z \in \{0, 1\}$]{\includegraphics[width=3.2in]{./images02/new-images/z_0_1.png}}

  \subfloat[$z \in \{0, 0.5, 1, 1.5, 3, 4.5, 6\}$]{\includegraphics[width=3.2in]{./images02/new-images/iter_z_all_2.png}}

  \caption{(a) and (b) show the behavior of the previous Figure, now executed with 6 iterative reads. What we observe clearly is that with a single read, as the distance grows, the system only `cleans' towards the orthoghonal distance 500 after a number of iterative readings.}
  \label{fig:murillo-generalization-experiments-6reads}
\end{figure}

Physicist Paulo Murilo observed that the models of Kanerva-read ($z=1$) and Chada-read ($z=0$) were simple variations of the exponent $z$, which suggests experimenting with different values. The results, however, have not yielded performance improvements.  Though for $z \leq 1$ results are comparable to $z=1$, for $z>1$, the system shows a clear deterioration, with a smaller distance to convergence and higher divergence at large-distance reads. This is shown in Figures \ref{fig:murillo-generalization-experiments} and \ref{fig:murillo-generalization-experiments-6reads}.

@@

\begin{figure}[h!]
  \centering
  \subfloat[$w_1(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-global.jpeg}}

  \subfloat[$w_1(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-to-500.jpeg}}

  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-stepwise.jpeg}}

  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius. (a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
  \label{fig:info-theory-hypothesis}
\end{figure}

\begin{figure}[h!]
  \centering
  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-a-sum.jpeg}}

  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-b-sum-range.jpeg}}

  \subfloat[stepwise $\left \lfloor{w_2(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-c-sum-stepwise.jpeg}}

  \caption{SOON TO BE DEPRECATED.  Shannon write operation:  Computing the sum of low-likelihood signals. (a) entirety of the space; (b) region of interest; (c) Fast integer computation through a stepwise function. }
  \label{fig:info-theory-sum-hypothesis}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \subfloat[$w_2(d), d \in \{ 1, 2, ..., n\}.$]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon.jpeg}}

  \subfloat[$w_2(d)$ for the desired range. ]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-region.jpeg}}

  \subfloat[stepwise $\left \lfloor{w_1(d)}\right \rfloor$ for fast integer computation.]{\includegraphics[width=3.2in]{./images02/new-images/Info-theory-shannon-stepwise.jpeg}}

  \caption{Shannon write operation:  Computing the amount of information of a signal to each hard location in its access radius. (a) entirety of the space; (b) region of interest; (c) Fast integer computation is possible through a stepwise function.}
  \label{fig:info-theory-hypothesis}
\end{figure}

My advisor, Alexandre Linhares, has proposed another read operation: an information-theoretical weighted reading. In it, the sum of the counter's value is weighted based on the distance between each hard-location's address and the reading address. The logic behind it is to vary the importance of each hard-location inside the circle.  It is only natural that one encodes an item in closer hard locations with a stronger signal, and a natural candidate for this signal function is the amount of information contained in the distance between the item and each hard location.  Closer hard locations have lower probabilies and therefore should encode more information.

@@

The initial results of this \emph{Shannon write} operation can be seen in Figure \ref{fig:info-theory-experiments} and seem promising. It seems that the critical distance increases by a number of bits.  Note that 10 additional bits imply an attractor $2^{10}$ of the size of the original. Another point to keep in mind is that, since the modulus of the vectors are not uniform in this approach, that the shape of the attractor may have asymmetries.

But, qualitatively, they seem disappointing.  HEATMAPFIGs

@@

Kanerva suggests, in his book, the use of 1,000 dimensions and 1,000,000 hard locations.  More recently, he suggested the use of 10,000 dimensions, and on personal discussions suggested that this should be a minimum; as he has been concerned in latent semantic analysis and seems to be the proper scale in that application.

@@

$\forall b \in \{ 0,1\} ^n$, we want an algorithm $A$ that yields the particular list of hard locations for $b$ and all hard locations respect the desired properties of the memory.  A reduction from measuring the distance to $10^6$ hard locations to a computation of $~ 10^3$ hard locations might yield astonishing performance gains, depending, of course, on our optimistic assumptions concerning existence and complexity of such algorithm.

At large scales of computing, the very ability to perform some experiments is a function of sheer performance. The horizon of experiments --- and possibly of knowledge --- expands \emph{as a function of computational demands}. More on this in my closing words.

\section{Seeing farther}

Let us revisit, in this concluding thoughts, the emphasis employed over speed of computation.  At first sight, that might seem like a typical objective of efficiency in computer science. But we are not only interested in the computer sciece effects here --- the ambition is different. More important than this `computer-sciency' goal, i.e., a beautiful, clean, efficient algorithm with the primary effect of enhanced speed, however, is the secondary effect on the sociology of science:  \emph{We can see farther}.

We have generated a Docker image, which makes it even easier to explore the framework. After running the container, a Jupyter Notebook is available with sdm-framework and other tools already installed.

All the simulations and graphics generated in this thesis are promptly available to be re-executed and explored by those interested. We invite readers to take a look and explore a little bit.
